---
title: "Problem Set 4"
author: "Sawera Dhaliwal (24324148)"
date: "3/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Part 1: Enumeration and Description
##### Load Packages
library(kableExtra)
library(knitr)
library(tidyverse)
library(vegan)

##### Load Community Data
community_data = data.frame(
  number = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16),
  name = c("rigo", "lego", "drop", "bear", "skit", "mnm", "mike","macrophage","swirl","watermelon","kisses","snake","bottle","hock","port","fish"),
  characteristics=c("wine gum","hard lego all sizes","gum drop","gummy bear","skittle","m and m chocolates","mike and ike","macrophage shaped gummy","swirl drop","watermelon gummy","kisses chocolate","long red gummy","gummy bottle","transluscent oval","translucent diamond","red green fish"),
  occurences = c(7,18,24,74,197,218,199,6,3,1,16,13,3,1,1,1)
)

##### Community Data Table
```{r}
community_data %>% 
  kable("html") %>%
  kable_styling(bootstrap_options = "striped", font_size = 10, full_width = F)
``` 

### Part 2: Collector’s Curve
(!!!)

### Part 3: Diversity Estimates (Alpha Diversity)

##### Simpson Reciprocal Index For Community Sample 
```{r}
species1 = 7/782
species2 = 18/782
species3 = 24/782
species4 = 74/782
species5 = 197/782
species6 = 218/782
species7 = 199/782
species8 = 6/782
species9 = 3/782
species10 = 1/782
species11 = 16/782
species12 = 13/782
species13 = 3/782
species14 = 1/782
species15 = 1/782
species16 = 1/782

1 / (species1^2 + species2^2 + species3^2 + species4^2 + species5^2 + species6^2 + species7^2 + species8^2 + species9^2 + species10^2 + species11^2 + species12^2 + species13^2 + species14^2 + species15^2 + species16^2)
```

##### Chao1 Richness Estimator
```{r}
16 + 4^2/(2*12)
``` 

### Part 4: Alpha-Diversity Functions in R
##### Remove Columns and Transpose Data
```{r}
community_data_diversity = 
  community_data %>% 
  select(name, occurences) %>% 
  spread(name, occurences)

community_data_diversity
```

##### Calculate Simpson Reciprocal Index
```{r}
diversity(community_data_diversity, index="invsimpson")
```

##### Calculate Chao1 Richness Estimator
```{r}
specpool(community_data_diversity)
```

(!!!) "for your sample...."

### Part 5: Concluding Activity 
(!!!)