---
title: "Problem Set 4"
author: "Sawera Dhaliwal (24324148)"
date: "3/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Part 1: Enumeration and Description
##### Load Packages
library(kableExtra)
library(knitr)
library(tidyverse)
library(vegan)

##### Load Community Data
community_data = data.frame(
  number = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16),
  name = c("rigo", "lego", "drop", "bear", "skit", "mnm", "mike","macrophage","swirl","watermelon","kisses","snake","bottle","hock","port","fish"),
  characteristics=c("wine gum","hard lego all sizes","gum drop","gummy bear","skittle","m and m chocolates","mike and ike","macrophage shaped gummy","swirl drop","watermelon gummy","kisses chocolate","long red gummy","gummy bottle","transluscent oval","translucent diamond","red green fish"),
  occurences = c(7,18,24,74,197,218,199,6,3,1,16,13,3,1,1,1)
)

##### Load Collector Data
collector_data = data.frame(
  number = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16),
  name = c("rigo", "lego", "drop", "bear", "skit", "mnm", "mike","macrophage","swirl","watermelon","kisses","snake","bottle","hock","port","fish"),
  characteristics=c("wine gum","hard lego all sizes","gum drop","gummy bear","skittle","m and m chocolates","mike and ike","macrophage shaped gummy","swirl drop","watermelon gummy","kisses chocolate","long red gummy","gummy bottle","transluscent oval","translucent diamond","red green fish"),
  occurences = c(2,2,8,17,27,41,27,2,2,0,5,3,0,0,0,0)
)

##### Community Data Table
```{r}
community_data %>% 
  kable("html") %>%
  kable_styling(bootstrap_options = "striped", font_size = 10, full_width = F)
``` 

##### Collector Data Table 
```{r}
collector_data %>% 
  kable("html") %>%
  kable_styling(bootstrap_options = "striped", font_size = 10, full_width = F)
```

##### Conclusion
The sample taken appears to be fairly representative of the community. Species appearing in high frequency remain prevalent in the sample. The sample does not contain five of the sixteen species identified in the community. This is not alarming, since the five species contained a small number of individuals to begin with. 

### Part 2: Collector’s Curve
##### Random Collection of Cells 
collector_sample_data = data.frame(
  x = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98,99,100,101,102,103,104,105,106,107,108,109,110,111,112,113,114,115,116,117,118,119,120,121,123,124,125,126,127,128,129,130,131,132,133,134,135,136),
  y = c(1,2,3,3,4,4,4,5,6,6,6,6,6,6,6,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,8,8,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,9,10,10,10,10,10,10,10,10,10,10,10,10,10,10,10,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11,11)
)

##### Creating the Plot 
```{r}
ggplot(collector_sample_data, aes(x=x, y=y)) +
  geom_point() +
  geom_smooth() +
  labs(x="Cumulative number of individuals classified", y="Cumulative number of species observed")
```

##### Conclusion
The curve flattens out entirely after the 79th cell collected. From the shape of the collector's curve, we are led to believe that the sampling was complete and representative of our community. 

### Part 3: Diversity Estimates (Alpha Diversity)
##### Simpson Reciprocal Index For Community Sample 
```{r}
species1 = 7/782
species2 = 18/782
species3 = 24/782
species4 = 74/782
species5 = 197/782
species6 = 218/782
species7 = 199/782
species8 = 6/782
species9 = 3/782
species10 = 1/782
species11 = 16/782
species12 = 13/782
species13 = 3/782
species14 = 1/782
species15 = 1/782
species16 = 1/782

1 / (species1^2 + species2^2 + species3^2 + species4^2 + species5^2 + species6^2 + species7^2 + species8^2 + species9^2 + species10^2 + species11^2 + species12^2 + species13^2 + species14^2 + species15^2 + species16^2)
```

##### Simpson Reciprocal Index For Collector's Sample 
```{r}
species1 = 2/136
species2 = 2/136
species3 = 8/136
species4 = 17/136
species5 = 27/136
species6 = 41/136
species7 = 27/136
species8 = 2/136
species9 = 2/136
species10 = 5/136
species11 = 3/136

1 / (species1^2 + species2^2 + species3^2 + species4^2 + species5^2 + species6^2 + species7^2 + species8^2 + species9^2 + species10^2 + species11^2)
```

##### Chao1 Estimate for Community Sample 
```{r}
16 + 4^2/(2*12)
``` 

##### Chao1 Estimate for Collector's Sample 
```{r}
11 + 0^2/(2*11)
``` 

##### Conclusion
The simspon REciprocal Index is 4.6 for the community and 5.2 for the sample.
The chao1 estimate is 16.7 for the community and 16 for the sample

### Part 4: Alpha-Diversity Functions in R
##### Remove Columns and Transpose Data for Community Data
```{r}
community_data_diversity = 
  community_data %>% 
  select(name, occurences) %>% 
  spread(name, occurences)

community_data_diversity
```

##### Remove Columns and Transpose Data for Collector's Data
```{r}
collector_data_diversity = 
  collector_data %>% 
  select(name, occurences) %>% 
  spread(name, occurences)

collector_data_diversity
```

##### Simpson Reciprocal Index for Community Data
```{r}
diversity(community_data_diversity, index="invsimpson")
```

##### Simpson Reciprocal Index for Collector Data
```{r}
diversity(collector_data_diversity, index="invsimpson")
```

##### Chao1 for Community Data
```{r}
specpool(community_data_diversity)
```

##### Chao1 for Collector Data
```{r}
specpool(collector_data_diversity)
```
##### Conclusion
The Simpson Reciprocal Indices for the community and collector's samples are 4.6 and 5.2 recpectively. The chao1 estimates for the community and collector's samples are 16 and 11 recpectively. This is almost entirely consistent with my previous calculations. The only discrepancy being a calculated Simpson value of 16.7 and a R function determined value of 16 for the community sample. 

### Part 5: Concluding Activity 
* How does the measure of diversity depend on the definition of species in your samples?  
For both the Simpson Reciprocal Index and Chao1 Richness Estimator methods, if you increase the number of defined species in your community, the estimated diversity will increase. 

* Can you think of alternative ways to cluster or bin your data that might change the observed number of species?  
For this assignment, my clusters were mostly based off of shape. The biggest change I could make, would be to cluster based off of both shape and color. This would increase the number of observed species and, consequently, the estimated diversity of my community. 

* How might different sequencing technologies influence observed diversity in a sample?  
I'm assuming different sequencing technologies (software) have unique built in species definition criteria. So, for example, one program may classify my sample based solely off of shape and the other both shape and color. As discussed above, this would result in different diversity estimates. Of course, this points towards the importance of finding a universal species definition, which would make comparison amongst studies and data consolidation much more effective. 